BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets

### 2024-07-11 manuscript/manuscript_APPS.rmd
adding sresults

@@

BarnebyLives took roughly `r minutemen(sum( time_trials[time_trials$Group=='Local', 'elapsed']))` minutes (`r sum( time_trials[time_trials$Group=='Local', 'elapsed'])`s) to run all local steps, and roughly `r minutemen(time_trials[time_trials$Function=='powo_searcher', 'elapsed'])` minutes (`r time_trials[time_trials$Function=='powo_searcher', 'elapsed']`s) to search Plants of the World Online, and `r time_trials[time_trials$Function=='directions_grabber', 'elapsed']`s to search Google Maps and write directions to sites. 
Most of the local run time is attributable to the spatial (spatial: `r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`s), and taxonomic operations (`r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`s), style: `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`s. 
The spell check operation of the scientific name accounted for nearly all of the time (`r sum(tt_local[tt_local$Function=='spell_check', 'elapsed'])`s) spent performing local taxonomic operations.
The generation of labels consumed around `r minutemen(time_label_gen)` minutes (`r time_label_gen`s) for the rendering, `r time_4perpage`s to combine individual labels four per single sheet of landscape orientated paper, and `r time_final`s to combine the `r length(list.files('labels/processed'))` sheets to a single Portable Document Format (PDF).  
The total computer run time for processing these XXX specimens was XX minutes.

@@

epithet_diff <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(Epithet != POW_Epithet, !Full_name %in% genus_mismatch$Full_name) %>% 
  select(Collection_number, Full_name, Epithet, POW_Genus, POW_Epithet, POW_Infraspecies) %>% 
  distinct(Full_name, .keep_all = T) %>%  # do we record repeat collections ?
  mutate(Situation = 
           case_when(
             Collection_number %in% c(
               '1217', '1251', '1283', '1311', '1337', '1391', '1450', '1576', 
               '1594', '1745', '2043', '2086', '2108', '2109', '2120', '2127', '2130',
               '2153', '2199', '2227', '2242', '2256', '2289', '2370', '2379', 
               '2393', '2415', '2511', '2573', '2648', '2660', '2687', '2690',
               '2710', '2715', '2750', '2767', '2818')
             ~ 'Spelling',
             Collection_number %in% c('1207') ~ 'Binomial_Reversed', 
             Collection_number %in% c('1588', '2445', '2777', '1337', '2412', '2748', 
                                      '2652', '2805', '2194', '2040', '2039', '2403',
                                      '2328', '2076') ~ 'Outdated',
             Collection_number %in% c('2644') ~ 'Incorrect',
             Collection_number %in% c('2326', '2038', '1424', '2618',
                                      '2665', '2297', '2343') ~ 'Internal_Error',
             Collection_number %in% c('2110') ~ 'Preferred',
             Collection_number %in% c('2480', '2561', '2727') ~ 'Not Found'

           )
  )

@@

Even on data which had been manually cleaned and error-checked by a human several times BarnebyLives was able to reduce transcription errors, identify typos, make nomenclature suggestions, and reformat text elements for downstream use. While no families were misspelled, BL made `r nrow(family_incongruence)` suggestions on naming, `r sum(family_incongruence$Situation=='Typo', na.rm = T)` manually entered typos were found, it identified `r sum(family_incongruence$Situation=='Submitter_Incorrect', na.rm = T)` instances where an incorrect family was entered, and `r sum(family_incongruence$Situation=='Outdated', na.rm = T)` instances of an outdated circumscription applied. BL flagged `r sum(family_incongruence$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy, and flagged `r sum(family_incongruence$Situation=='Internal_Error', na.rm = T)` records in error.

BL identified `r nrow(genus_mismatch)` discrepancies at the level of genus between user submitted and processed data. 
In `r sum(genus_mismatch$Situation=='Outdated')` of these instances the user supplied an outdated name instance of an outdated circumscription applied (`r nrow(unique(genus_mismatch[genus_mismatch$Situation=='Outdated', 'Genus']))` genera total). BL flagged `r sum(genus_mismatch$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy (`r nrow(unique(genus_mismatch[genus_mismatch$Situation=='Preferred', 'Genus']))` genera total), and flagged `r sum(genus_mismatch$Situation=='Internal_Error', na.rm = T)` record in error.

BL flagged 75 species

>>>>

n_genera <- data %>% 
  separate(Scientific_name, into = c('Genus', 'Epithet') , remove = FALSE, extra = 'drop') %>% 
  group_by(Genus) %>% 
  count() %>% 
  nrow() # 292 Genera

@@

BarnebyLives took roughly `r minutemen(sum( time_trials[time_trials$Group=='Local', 'elapsed']))` minutes (`r sum( time_trials[time_trials$Group=='Local', 'elapsed'])`s) to run all local steps, and roughly `r minutemen(time_trials[time_trials$Function=='powo_searcher', 'elapsed'])` minutes (`r time_trials[time_trials$Function=='powo_searcher', 'elapsed']`s) to search Plants of the World Online, and `r time_trials[time_trials$Function=='directions_grabber', 'elapsed']`s to search Google Maps and write directions to sites. 
Most of the local run time is attributable to the spatial (spatial: `r sum(tt_local[tt_local$Module=='Geospatial', 'elapsed'])`s), and taxonomic operations (`r sum(tt_local[tt_local$Module=='Taxonomic', 'elapsed'])`s), style: `r sum(tt_local[tt_local$Module=='Style', 'elapsed'])`s. 
The spell check operation of the scientific name accounted for nearly all of the time (`r sum(tt_local[tt_local$Function=='spell_check', 'elapsed'])`s) spent performing local taxonomic operations.
The generation of labels consumed around `r minutemen(time_label_gen)` minutes (`r time_label_gen`s) for the rendering, `r time_4perpage`s to combine individual labels four per single sheet of landscape orientated paper, and `r time_final`s to combine the `r length(list.files('labels/processed'))` sheets to a single Portable Document Format (PDF).  
The total computer run time for processing these `r nrow(data)` specimens was `r minutemen(sum(time_trials$elapsed))` minutes.

@@

epithet_diff <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(Epithet != POW_Epithet, !Full_name %in% genus_mismatch$Full_name) %>% 
  select(Collection_number, Full_name, Epithet, POW_Genus, POW_Epithet, POW_Infraspecies) %>% 
  distinct(Full_name, .keep_all = T) %>%  # do we record repeat collections ?
  mutate(Situation = 
           case_when(
             Collection_number %in% c(
               '1217', '1251', '1283', '1311', '1337', '1391', '1450', '1576', 
               '1594', '1745', '2043', '2086', '2108', '2109', '2120', '2127', '2130',
               '2153', '2199', '2227', '2242', '2256', '2289', '2297', '2343', '2370', '2379', 
               '2393', '2415', '2480', '2511', '2561', '2573', '2648', '2660', 
               '2687', '2690', '2710', '2715', '2727', '2750', '2767', '2818')
             ~ 'Spelling',
             Collection_number %in% c(
               '1588', '1619', '2445', '2777', '1337', '2412', '2487', '2748', '2652',
               '2805', '2194', '2040', '2039', '2403', '2328', '2076') 
             ~ 'Outdated', 
             Collection_number %in% c('2644') ~ 'Incorrect',
             Collection_number %in% c(
               '1207', '2326', '2038', '1424', '2618', '2622', '2665', '2711', '2713') 
             ~ 'Internal_Error',
             Collection_number %in% c('2110', '2540') ~ 'Preferred'
           )
  )

@@

Even on data which had been manually cleaned and error-checked by a human several times BarnebyLives was able to reduce transcription errors, identify typos, make nomenclature suggestions, and reformat text elements for downstream use. 
While none of the `r n_families` families were misspelled, BarnebyLives made `r nrow(family_incongruence)` suggestions on naming, identified `r sum(family_incongruence$Situation=='Typo', na.rm = T)` typos, identified `r sum(family_incongruence$Situation=='Submitter_Incorrect', na.rm = T)` instances where an incorrect family was entered, and `r sum(family_incongruence$Situation=='Outdated', na.rm = T)` instances of an outdated circumscription applied. 
At the level of family BarnebyLives flagged `r sum(family_incongruence$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy, and flagged `r sum(family_incongruence$Situation=='Internal_Error', na.rm = T)` records in error.

In the `r n_genera` genera analysed  BarnebyLives identified `r nrow(genus_mismatch)` discrepancies at the level of genus between user submitted and processed data. 
In `r sum(genus_mismatch$Situation=='Outdated')` of these instances the user supplied an outdated name (`r nrow(unique(genus_mismatch[genus_mismatch$Situation=='Outdated', 'Genus']))` unique genera). 
BL flagged `r sum(genus_mismatch$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy (`r nrow(unique(genus_mismatch[genus_mismatch$Situation=='Preferred', 'Genus']))` genera total), and flagged `r sum(genus_mismatch$Situation=='Internal_Error', na.rm = T)` records in error.

Of the `r nrow(data)` species analysed (`r n_spp` distinct species) BarnebyLives flagged `r nrow(epithet_diff)` records. 
BL `r sum(epithet_diff$Situation=='Spelling')` detected instances of misspelled epithets (`r sum(epithet_diff$Situation=='Spelling')` unique species). 
In `r sum(epithet_diff$Situation=='Outdated')` of these instances the user supplied an outdated name (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Outdated', 'Full_name']))` unique species). 
BL flagged `r sum(epithet_diff$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Preferred', 'Full_name']))` unique species), and flagged `r sum(epithet_diff$Situation=='Internal_Error', na.rm = T)` records in error. 
The final record was an egregious error where the order of the specific epithet and the genus name.
